Formalizing Stack Safety as a Security Property

### 2023-02-04 submission5/core.tex

The term {\em stack safety} is used for a variety of compiler,
run-time, and hardware mechanisms for protecting stack memory. Unlike
``the heap,'' the ISA-level stack does not correspond to a single
high-level language concept: different compilers use it in different
ways to support procedural and functional abstraction mechanisms from
a wide range of languages.  This protean nature makes it difficult to
nail down what it means to correctly enforce stack safety.

@@

This formulation is motivated by a particular class of enforcement
mechanisms, the ``lazy'' stack safety micro-policies studied by
Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, which permit
functions to write into one another's frames, but which taint the
changed locations so that the frame's owner cannot access it. No
existing characterization of stack safety captures this style of
safety. We capture it here by stating our properties in terms of the
observable behavior of the system.

% The stack plays a central role in realizing large number of
% high-level language features, some of which are typically excluded
% from discussions of stack safety. 
Our properties go further than previous formal definitions of stack
safety, supporting caller- and callee-saved registers, arguments
passed on the stack, \ifexceptions tail-call elimination, and
exceptions.  \else and tail-call elimination. \fi 
% \bcp{Out of all of
%   these, only exceptions are high-level language features!}
They are modular by design so that they can be further extended to
other features.

We validate our properties by using them to distinguish between
correct and incorrect implementations of Roessler and DeHon's
micro-policies using property-based random testing. Our test harness
successfully identifies several broken variants, including Roessler
and DeHon's lazy policy; a repaired version of their policy does pass
our tests.

@@

{\em Functions} in high-level languages, and related abstractions such
as subroutines, procedures, methods, etc., are units of computation
that call one another to modularly realize larger computations.
%
At a low level, each function activation manages its own data---local
variables, etc.---as well as information about the \emph{caller} to
which it must return.
%
The \emph{call stack} is the fundamental data structure used to
implement functions, aided by an ABI that defines how registers are shared
between activations.

From a security perspective, attacks on the call stack
are attacks on the function abstraction itself.
%
Indeed, the stack is an ancient~\cite{phrack96:smashingthestack} and
perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
  DBLP:conf/sp/SzekeresPWS13,
  DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
target for low-level attacks, sometimes involving control-flow
hijacking via corrupting the return address, sometimes memory corruption
more generally.  
%
%The most recent release of the authoritative CWE Top 25
%Most Dangerous Software Weaknesses~\cite{
%list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
%shows that general classes of vulnerabilities closely related to the stack
%% \rb{(though not exclusively)}
%consistently rank at and near the top of the list, e.g., \#1 Out-of-bounds Write, \#5 Out-of-bounds Read, and \#7 Use After Free.
%
%Moreover, many other types of faults involve some kind of control or data flow
%using data stored in the stack 
The variety in attacks on the stack is mirrored in the range of
software and hardware protections that aim to prevent them,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18,Gollapudi+23}.

Enforcement mechanisms can be brittle, successfully eliminating one
attack while leaving room for others. To avoid an endless game of
whack-a-mole, we seek formal properties of safe behavior that can be
proven or at least rigorously tested. Such properties can be used as
the specification against which enforcement can be validated; even
enforcement mechanisms that do not fulfill a property benefit from the
ability to articulate why and when they may fail.

Of the mechanisms listed above, many are fundamentally ill suited for
offering formal guarantees: may impede attackers, but do not provide
universal protection. Shadow stacks, for instance, aim to ``restrict
the flexibility available in creating gadget chains''
\cite{Shanbhogue+19}, not to categorically rule out attacks. Other
mechanisms, such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
integrity\cite{Kuznetsov+14}, do aim for stronger guarantees, but even
these do not make formal promises.  Indeed, to our knowledge, the sole
line of work making a formal claim to protect stack safety is the
secure calling conventions of Skorstengaard et
al.~\cite{SkorstengaardSTKJFP} and Georges et
al.~\cite{Georges22:TempsDesCerises}.

Other mechanisms besides these should also be amenable to strong
formal guarantees.  In particular, Roessler and DeHon
\cite{DBLP:conf/sp/RoesslerD18} have presented an array of tag-based
micro-policies~\cite{pump_oakland2015} for stack safety that offer
strong guarantees.  Their most realistic micro-policy, called Lazy
Tagging and Clearing (\(LTC\)), makes an interesting performance
trade-off: it allows function activations to improperly write into one
another's stack frames, but ensures that the owner of the corrupted
memory cannot access it afterward.

Under this policy, one function activation {\em can} corrupt another's
memory---just not in ways that do not affect observable behavior.
Therefore, \(LTC\) would not fulfill Georges et al.'s property
(appropriately adapted to the tagged setting).  But \(LTC\) does
arguably enforce stack safety.  A looser, more observational
definition of stack safety is needed to fit this situation.

We propose here a formal characterization of stack safety, based on
the intuition of protecting function activations from each other.
\bcp{Aren't other formal definitions based on the same intuition?}
We use the tools of language-based
security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the
{\em well-bracketed control flow} (\(\wbcf\)) property articulated by
Skorstengaard et al.~\cite{SkorstengaardSTKJFP}.
While our properties are motivated by the desire to specify \(LTC\) precisely, 
they are stated abstractly in hopes that they may also apply
to other enforcement mechanisms.

We find, via property-based random testing, that \(LTC\) is flawed in
a way that undermines both integrity and confidentiality; after
correcting this flaw, the repaired \(LTC\) satisfies all our
properties.  Moreover, we show that our properties can be extended to
describe calling conventions with features such as callee-save
registers, argument-passing on the stack, and tail-call
elimination. We modify \(LTC\) to protect these features and apply
random testing to validate this extended protection mechanism against
the extended properties.

@@

We show that we can extend our system to new features by first introducing a simple
example model, and then moving to a more realistic model with multiple new features:
callee-save registers, argument passing on the stack, and tail-call elimination, simply
by refining a few definitions in (1). This becomes the model that we validate.
In \cref{app:ptr}, we extend the model further with arbitrary sharing of memory-safe
pointers. Our criteria in (3) cover a total of five core requirements:
\(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
They are related to one another as seen in
\cref{fig:propertydiagram}:\leo{is this supposed to be \cref{tab:props}?} \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
\(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
according to the needs of a system.

In particular, many enforcement techniques focus purely on
well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
address, and shadow stacks with protection (e.g. Return Address Defender \cite{Chiueh2001RAD})
to enforce it completely. Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Interestingly, Skorstengaard et al. \cite{SkorstengaardSTKJFP}
describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by 
Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
When checking if a preliminary version of these properties applied to
the Cerise calling convention, they noted that the Skostengaard et al. calling convention
did not enforce callee confidentiality, and made sure that their version would
\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
This demonstrates the benefit of our choice to explicitly state properties in security
terms: specifying security mechanisms is hard, and when the specification takes the
form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
security requirement.

@@

\begin{itemize}
\item We give a novel characterization of stack safety as a collection
of properties: confidentiality and integrity for callee and caller,
plus well-bracketed control-flow. Different subsets of these
properties characterize the security guarantees offered by mechanisms
that support only some aspects of stack safety\bcp{where do we show
  this?}.  The properties they are parameterized over a notion of
external observation, allowing them to characterize lazy enforcement
mechanisms.
\item We instantiate\bcp{extend??} these definitions in a realistic
setting with features such as\bcp{``such as''? are there others?}
argument passing on the stack, callee-saves registers, and tail-call
elimination.\bcp{How about: We extend these core definitions to
  describe a realistic setting with argument passing on the stack,
  callee-saves registers, and tail-call elimination.}  Our model is
modular enough that handling these features is straightforward.
\bcp{Throughout, the way we talk about the ``core model'' and its
  intantiation / extension / enrichment with realistic features feels
  quite weasily.}
\item We validate tag-based stack-safety enforcement via
property-based randomize testing. We find that \emph{Lazy Tagging and
  Clearing} falls short of ensuring stack safety, propose a fix, and
validate it.
\end{itemize}

@@

\section{Building a Property}
\label{sec:ideas}

Our properties follow a three-part structure that we term a {\em security semantics}.
First we define the {\em security-relevant operations} that we support, and an extra
piece of state called a {\em security context}, which updates alongside each step
of execution. The security context gives each function activation a {\em view},
mapping each element to a {\em security class}.

Then, we define our high-level security requirements---integrity, confidentiality,
and well-bracketed control flow---in terms of the security classes of state
elements. These take the form of predicates on states and contexts that use variants,
assertions about future return states, and comparison of event traces to capture the
desired behavior without reference to implementation details.

Separately, in order to apply a property to a specific machine, we do need to consider
implementation details, by determining how the machine's execution translates into
security relevant operations. But this step is transparent to the properties themselves.

\apt{Something about how the machine can incorporate enforcement, with examples
(tags, capabilities, SW only). }

\paragraph*{Security Semantics and Property Structure}

A security semantics extends a machine
with additional context about the identity of current and pending
functions (which act as security principals) and about the registers and memory they require
to be secure. This added context is purely notional;
it does not affect the real machine. The security context
evolves dynamically through the execution of security-relevant operations,
which include events like calls, returns, and frame manipulations.
Our security properties are phrased in terms of this context, often as predicates
on future states, e.g. of the form ``when control returns to the current function...'',
or as relations on traces of future execution (hyper-properties).

The security-relevant operations typically correspond to underling machine instructions,
but the correspondence may not be obvious just by inspecting the machine code.
For example, in the tagged RISC-V machine we use in our examples and tests,
calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
instructions might also be used for other things. An instantiation of an operation
might require some external mechanism for distinguishing which instructions are
security-relevant. The most obvious technique, which we use in our examples, is to add labels to
selected instructions, but the theory does not depend on this choice.

Whatever the mechanism for labeling transitions, we get a base
machine transition \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
an event (see below) and \(\bar{\psi}\) is a list of security-relevant operations.
%We then lift this into a transition between pairs of machine states and contexts,
%\((\mach,\context) \stepstounder{\overline{\psi},\obs} (\mach', \context')\).
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. Those that we explore in detail in \cref{sec:example,sec:formal}
are shown in white: calls, returns, and the allocation and deallocation of local memory.
In \cref{sec:extensions}, we cover additional operations and and parameters 
to support various forms of memory sharing \ifexceptions , exceptions, \fi
and tail-call elimination; these are shown in light gray. These are also
the operations we test in \cref{sec:testing}.
We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
based on pointer provenance in \cref{app:ptr}, represented by the dark gray
operations, but we do not test this.

@@

\begin{table}
\begin{center}
  \begin{tabular}{| l | l |}
    \hline
    \(\psi \in \Psi\) & Parameters \\
    \hline
    \example \(\mathbf{call}\) & target address, argument registers \\
    \testing & stack arguments (base, offset \& size) \\
    \example \(\mathbf{return}\) & \\
    \example \(\mathbf{alloc}\) & offset \& size \\
    \testing & public flag \\
    \example \(\mathbf{dealloc}\) & offset \& size \\
    \testing \(\mathbf{tail call}\) & (same as for \(\mathbf{call}\)) \\
    \hline
    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
    \hline
    \theory \(\mathbf{promote}\) & register, offset \& size \\
    \theory \(\mathbf{propagate}\) & source register/address \\
    \theory & destination register/address \\
    \theory \(\mathbf{clear}\) & target register/address \\
    \hline
  \end{tabular}
\end{center}
\caption{Security-relevant operations and their parameters.}
  \label{tab:psi}
\end{table}

Our security semantics has similarities to the overlay semantics that have been proposed
to characterize stack safety in the past~\cite{SkorstengaardSTK}, but it does not restrict
the behavior of the underlying machine in any way. Rather, it tracks additional context
about the history of security-relevant operations, which inform the criterion a machine
must satisfy in order to correctly obey the properties.

The security principals in our account of stack safety are function
activations: the context tracks, for the active function and all pending
functions, their {\em view} of the security state of the system.

\paragraph*{Views}

@@

\paragraph*{Events}

It is common to specify the behavior of a system, including security properties, in
terms of traces of observable {\em events} rather than details of the machine state.
Lazy enforcement works because the deferred check catches errors before they appear
in the trace. The nature of events is a parameter of the machine, and can be specialized
to any notion of observable behavior.

\paragraph*{Variants and Non-interference}

To characterize situations in which values should be kept secret or prevented
from influencing future execution, we borrow the idea of \emph{variant} states
from the theory of non-interference. Non-interference is a way of expressing
the flow of information through the system. If we begin with a state \(\mach\)
and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
then execute both \(\mach\) and \(\nach\), any differences in the outputs they
produce leaks information about \(\component\).

@@

Finally, the core property definitions are given in \cref{tab:props},
arranged to show their commonalities and distinctions. Each definition gives a criterion
quantified over states \(s\) that immediately follow call steps.
If an execution includes a transition \(s' \stepstounder{\overline{\psi}} s\)
where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
of a call.
%Likewise, if \(\mathbf{tail call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
%\(s\) is the target of a tail call.
As a shorthand, we write that each property is defined
by a criterion that must hold ``for all call targets \(s\),'' or, in the case of \(\wbcf\),
``for all call steps \(s \stepstounder{} s'\).''

@@

The operation rule for a tail call is similar to that for a normal call.
We do not push the caller's view onto the stack,
but replace it outright. This means that a tail call does not increase the size of
the call stack, and therefore for purposes of our properties, all tail
calls will
be considered to return simultaneously when the eventual {\bf return} operation
pops the top of the stack.

Since the caller will not be returned to, it does not need integrity, but
it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
than \(\sealed\) to express this. In terms of \cref{tab:props}, we replace
``call'' with ``call or tail call'' in \(\clrc\), \(\clec\), and \(\clei\).

@@

The centrality of the functional abstraction and its security are behind the
many software and hardware mechanisms proposed for its protection
\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey in
Sections~\ref{sec:intro} and~\ref{sec:future}. Among these techniques, the micro-policies of Roessler and
DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
range of safeguards including lazy enforcement, and suitable for formulating
and validating strong security guarantees.

>>>>

The term {\em stack safety} is used for a variety of compiler,
run-time, and hardware mechanisms for protecting stack memory. Unlike
``the heap,'' the hardware stack does not correspond to a single
high-level language concept: different compilers use it in different
ways to support procedural and functional abstraction mechanisms from
a wide range of languages.  This protean nature makes it difficult to
nail down what it means to correctly enforce stack safety.

@@

This presentation is motivated by a particular class of enforcement
mechanisms, the ``lazy'' stack safety micro-policies studied by
Roessler and DeHon~\cite{DBLP:conf/sp/RoesslerD18}, which permit
functions to write into one another's frames, but which taint the
changed locations so that the frame's owner cannot access it. No
existing characterization of stack safety captures this style of
safety. We capture it here by stating our properties in terms of the
observable behavior of the system.

The hardware stack interacts with a large number of language features,
some of which are typically excluded from discussions of stack
safety. Our properties go further than previous proposals,
supporting caller- and callee-saved registers, arguments passed on the
stack, \ifexceptions tail-call elimination, and exceptions.  \else and
tail-call elimination. \fi They are modular by design so that they
can be further extended to other features.

We validate our properties by using them to distinguish between
correct and incorrect implementations of Roessler and DeHon's
micro-policies using property-based random testing. Our test harness
successfully identifies several broken variants, including Roessler
and DeHon's original lazy policy.  A fixed version of their policy
does pass our tests.

@@

Computer programming relies on functions (or subroutines, procedures, methods, etc.)
as a foundational abstraction. Functions are units of computation that
can call one another to realize larger computations in a modular fashion.
%
At a low level, a function activation manages its own data, such as local variables, as
well as information about its \emph{caller} function, to which it must return.
The \emph{(call) stack} is the fundamental data structure charged with realizing
this abstraction, aided by an ABI that defines how registers are shared between activations.
From a security perspective, attacks on the stack are really attacks
on the fundamental functional abstraction itself, and are therefore every bit as
pervasive.

Indeed, the call stack is an ancient~\cite{phrack96:smashingthestack} and
perennial~\cite{mitre-cwe,DBLP:conf/raid/VeendCB12,
  DBLP:conf/sp/SzekeresPWS13,
  DBLP:conf/sp/HuSACSL16,msrc-bluehat,chromium-security}
target for low-level attacks, frequently involving control-flow hijacking via corrupting
the return address, and memory corruption more generally.
%
%The most recent release of the authoritative CWE Top 25
%Most Dangerous Software Weaknesses~\cite{
%list\footnote{\url{https://cwe.mitre.org/top25/archive/2022/2022_cwe_top25.html}}
%shows that general classes of vulnerabilities closely related to the stack
%% \rb{(though not exclusively)}
%consistently rank at and near the top of the list, e.g., \#1 Out-of-bounds Write, \#5 Out-of-bounds Read, and \#7 Use After Free.
%
%Moreover, many other types of faults involve some kind of control or data flow
%using data stored in the stack 
The variety in attacks on the stack is mirrored by the
software and hardware protections that aim to prevent them,
%
including stack canaries~\cite{Cowan+98},
bounds checking~\cite{NagarakatteZMZ09,NagarakatteZMZ10,DeviettiBMZ08},
split stacks~\cite{Kuznetsov+14},
shadow stacks~\cite{Dang+15,Shanbhogue+19},
capabilities~\cite{Woodruff+14,Chisnall+15,SkorstengaardLocal,SkorstengaardSTKJFP,Georges22:TempsDesCerises},
and hardware tagging~\cite{DBLP:conf/sp/RoesslerD18,Gollapudi+23}.

Security mechanisms can be brittle, successfully eliminating one attack while leaving room for
other, similar attacks. To avoid this \leo{game of whack-a-mole}, we seek formal properties of safe behavior that can be proven
or tested. Such properties become the specification against which enforcement can be validated;
even enforcement mechanisms that do not fulfill a property benefit from the ability to articulate
\emph{why} they fail.

Of the mechanisms above, many are fundamentally ill-suited for offering formal guarantees,
as they aim to impede attackers but cannot provide universal protection. Shadow stacks,
for instance, aim to ``restrict the flexibility available in creating gadget chains''
\cite{Shanbhogue+19}, not to categorically
rule out attacks. Other mechanisms such as Softbound\cite{NagarakatteZMZ09} and Code-pointer
integrity\cite{Kuznetsov+14} are stronger, but most do not give formal guarantees.
Indeed, the secure calling conventions of Skorstengaard et al. \cite{SkorstengaardSTKJFP} and
Georges et al. \cite{Georges22:TempsDesCerises} is so far the sole line of work
to offer stack safety formalizations.

Among the mechanisms that should be amenable to making strong formal guarantees,
Roessler and DeHon \cite{DBLP:conf/sp/RoesslerD18} present an array of tag-based
micro-policies~\cite{pump_oakland2015} for stack safety.
Their most realistic micro-policy, ``Lazy Tagging and Clearing'' (\(LTC\)),
makes an interesting trade-off for performance: it allows function activations to write
into one another's stack frames, but ensures that the owner of the corrupted memory cannot access
it afterward. Because the memory does get corrupted, \(LTC\) would not fulfill an
adaptation of Georges et al.'s property to the tagged setting.

Neverthess, we believe that there is a useful sense in which \(LTC\) does (attempt to) enforce stack safety. 
While it does not always protect individual addresses on the stack, it tries to 
prevent one function activation from interfering with another's behavior---protecting 
the functional abstraction, which we argue is the real essence of stack safety.

In this work, we propose a formal characterization of stack safety based on the idea
of protecting each function activation from other activations.
We use the formal tools of language-based
security~\cite{sabelfeld2003language}, decomposing stack safety into a family of
properties describing
the {\em integrity} and {\em confidentiality} of the caller’s local state
and the callee's behavior during the callee's execution, plus the control-flow protection
of {\em well-bracketed control flow} (\(\wbcf\))~\cite{SkorstengaardSTKJFP}.
While our properties are motivated by a desire to specify \(LTC\), 
they are built from first principles in the hopes that they will be broadly applicable
to a range of enforcement mechanisms.

The structure of our properties enables them to be easily expanded to characterize systems with
sophisticated features such as callee-save registers, argument-passing on the stack,
and tailcall-elimination. We likewise extend \(LTC\) to protect these features, and
test the extended micro-policy against our property. We find that, apart from the extensions, \(LTC\)
is flawed in a way that undermines integrity and confidentiality, and we correct
this flaw.

@@

We show that we can extend our system to new features by first introducing a simple
example model, and then moving to a more realistic model with multiple new features:
callee-save registers, argument passing on the stack, and tailcall elimination, simply
by refining a few definitions in (1). This becomes the model that we validate.
In \cref{app:ptr}, we extend the model further with arbitrary sharing of memory-safe
pointers. Our criteria in (3) cover a total of five core requirements:
\(\wbcf\), caller integrity (\(\clri\)), caller confidentiality (\(\clrc\)),
callee integrity (\(\clei\)), and callee confidentiality (\(\clec\)).
They are related to one another as seen in
\cref{fig:propertydiagram}:\leo{is this supposed to be \cref{tab:props}?} \(\clei\) subsumes \(\clrc\) and \(\clec\) subsumes
\(\clri\), but otherwise the properties are orthogonal and can be taken piecemeal
according to the needs of a system.

In particular, many enforcement techniques focus purely on
well-bracketed control flow. Stack canaries aim to prevent certain attacks on the return
address, and shadow stacks with protection (e.g. Return Address Defender~\cite{Chiueh2001RAD})
to enforce it completely. Others combine this protection with some degree of memory protection,
chiefly focusing on integrity. Interestingly, Skorstengaard et al.~\cite{SkorstengaardSTKJFP}
describe their ``local state encapsulation'' in terms of integrity, but it is equivalent
to the combination of \(\clri\) and \(\clrc\). In fact, the follow-up work by 
Georges et al. \cite{Georges22:TempsDesCerises} was inspired by the properties presented in this paper to extend their formal discussion to include confidentiality.
When checking if a preliminary version of these properties applied to
the Cerise calling convention, they noted that the Skostengaard et al. calling convention
did not enforce callee confidentiality, and made sure that their version would
\cite{Georges22:personalcommunication}, in addition to building it into their formalism.
This demonstrates the benefit of our choice to explicitly state properties in security
terms: specifying security mechanisms is hard, and when the specification takes the
form of a ``correct by construction'' machine, it is easy to neglect a non-obvious
security requirement.

@@

\begin{itemize}
\item A novel characterization of stack safety as a collection of properties:
  confidentiality and integrity for callee and caller, and well-bracketed
  control-flow. These can characterize the security requirements of systems
  that support only some portion of stack safety, and are
  parameterized over a notion of external observation, so that they can characterize
  lazy enforcement.
\item An instantiation of these definitions in a realistic
  setting with features such as argument passing
  on the stack, callee-saves registers, and tail-call elimination.
  Our model is modular enough that handling these features is straightforward.
\item The validation, via property-based randomize testing, of tag-based
  stack-safety policies. We find that \emph{Lazy Tagging and Clearing}
  falls short of stack safety, and propose a fix.
\end{itemize}

@@

\section{Framework and Assumptions}
\label{sec:ideas}

Stack safety properties need to describe the behavior of machine code. but they naturally 
talk about function activations and stack contents---abstractions that
are (typically) not visible at machine level. To bridge this gap, 
our properties are defined in terms of a {\em security semantics} layered on top of
the standard execution semantics of the machine.  This semantics identifies certain
state transitions of the machine as {\em security-relevant operations}, which update
a notional {\em security context}.  This context consists of
an (abstract) stack of function activations, each associated with a {\em view} 
that maps each element of the machine state (memory location or register)
to a {\em security class} (active, sealed, etc.) which specifies how the activation
can access the element.
The action of the security-relevant operations on the context is defined by a set
of rules that depend on how the machine code makes use of the stack and registers
to implement the function abstraction.

We define stack safety properties in
terms of traces of {\em observable events} rather than internal machine states. 
This is particularly important for
specifying lazy enforcement, which works because the deferred check catches problems 
before they can affect the trace.
Events might include certain function calls (i.e., system calls performing I/O)
or writes to special addresses representing memory-mapped regions, etc.
%The nature of events is a parameter of the framework, and can be specialized
%to any notion of observable behavior.

Finally, we define our high-level security properties---integrity, confidentiality,
and well-bracketed control flow---in terms of the security classes of state
elements. These take the form of predicates on states and contexts that use variants,
assertions about future return states, and comparison of event traces to capture the
desired behavior without reference to internal implementation details.
We introduce these properties by example in \cref{sec:example}. In the remainder
of this section we describe the underlying framework in more detail.

Our security semantics has similarities to the overlay semantics proposed
by Skorstengaard et al.~\cite{SkorstengaardSTKJFP}, but it does not restrict
the behavior of the underlying machine in any way. Rather, it tracks additional context
about the history of security-relevant operations, which inform the criterion a machine
must satisfy in order to correctly obey the properties.\apt{Sharpen this comparison?}

%The security principals in our account of stack safety are function
%activations: the context tracks, for the active function and all pending
%functions, their {\em view} of the security state of the system.

\paragraph*{Machine model}
We assume a conventional processor ISA (e.g. RISC-V, x86-64, etc.), with registers including a program counter
and stack pointer, and a standard ABI.
We make no particular assumptions about the provenance of the machine code; in particular,
we do not assume the use of any particular compiler.
The machine may possibly be enhanced with security
enforcement mechanisms such as hardware tags~\cite{pump_hasp2014,Gollapudi+23} or capabilities~\cite{Woodruff+14}. If so,
we assume that the behavior of these mechanisms is incorporated into the basic
step semantics of the machine, but we distinguish the \emph{payload} part of
each state element---the part relevant to the behavior of the basic machine---from
any additional enforcement information (such as a tag).

\paragraph*{Security Semantics and Property Structure}
A security semantics extends a machine
with additional context about the identity of current and pending
functions (which act as security principals) and about the registers and memory they require
to be secure. This added context is purely notional;
it does not affect the real machine. The security context
evolves dynamically through the execution of security-relevant operations,
which include events like calls, returns, and frame manipulations.
Our security properties are phrased in terms of this context, often as predicates
on future states, e.g. of the form ``when control returns to the current function...'',
or as relations on traces of future execution (hyper-properties).

The security-relevant operations typically correspond to underling machine instructions,
but the correspondence may not be obvious just by inspecting the machine code.
For example, in the tagged RISC-V machine we use in our examples and tests,
calls and returns are conventionally performed by {\tt jal} (``jump-and-link'')
and {\tt jalr} (``jump-and-link-register'') instructions, respectively, but these
instructions might also be used for other things. An instantiation of an operation
might require some external mechanism for distinguishing which instructions are
security-relevant. The most obvious technique, which we use in our examples, is to add labels to
selected instructions, but the theory does not depend on this choice.
Whatever the mechanism for labeling transitions, we end up with an annotated version of the 
machine transition function \(\mach \xrightarrow{\bar{\psi}, \obs} \mach'\), where \(\obs\) is
an event and \(\bar{\psi}\) is a list of security-relevant operations.

We then lift this into a transition between pairs of machine states and contexts,
by applying a set of operation-specific rules to transform each state and context to
a new context, in parallel with the ordinary transition on states.
The most important rules describe call and return operations.
A call pushes a new activation view onto the context stack and changes the class of the
caller's data to protect it from the new callee; a return reverses these steps.
Other operations can serve to signal how parts of the stack frame are being used to store
or share data, and their corresponding rules alter the classes of different
state elements accordingly.

Exactly which operations and rules are needed depends on
what code features we wish to support. 
The set of security-relevant operations (\(\Psi\)) covered in this paper is given in
\cref{tab:psi}. Those that we explore in detail in \cref{sec:example,sec:formal}
are shown in white: calls, returns, and the allocation and deallocation of local memory.
In \cref{sec:extensions}, we cover additional operations and and parameters 
to support various forms of memory sharing \ifexceptions , exceptions, \fi
and tail-call elimination; these are shown in light gray. These are also
the operations we test in \cref{sec:testing}.
We give a proof-of-concept formalization of a more sophisticated, heap-like sharing model
based on pointer provenance in \cref{app:ptr}, represented by the dark gray
operations, but we do not test this.

@@

\begin{table}
\begin{center}
  \begin{tabular}{| l | l |}
    \hline
    \(\psi \in \Psi\) & Parameters \\
    \hline
    \example \(\mathbf{call}\) & target address, argument registers \\
    \testing & stack arguments (base, offset \& size) \\
    \example \(\mathbf{return}\) & \\
    \example \(\mathbf{alloc}\) & offset \& size \\
    \testing & public flag \\
    \example \(\mathbf{dealloc}\) & offset \& size \\
    \testing \(\mathbf{tailcall}\) & (same as for \(\mathbf{call}\)) \\
    \hline
    \multicolumn{2}{|c|}{{\it Pointer provenance operations}} \\
    \hline
    \theory \(\mathbf{promote}\) & register, offset \& size \\
    \theory \(\mathbf{propagate}\) & source register/address \\
    \theory & destination register/address \\
    \theory \(\mathbf{clear}\) & target register/address \\
    \hline
  \end{tabular}
\end{center}
\caption{Security-relevant operations and their parameters.}
  \label{tab:psi}
\end{table}

\paragraph*{Views and security classes}

@@

%% \paragraph*{Variants and Non-interference}

%% To characterize situations in which values should be kept secret or prevented
%% from influencing future execution, we borrow the idea of \emph{variant} states
%% from the theory of non-interference. Non-interference is a way of expressing
%% the flow of information through the system. If we begin with a state \(\mach\)
%% and modify it to \(\nach = \mach[\component \mapsto v]\) for arbitrary \(v\),
%% then execute both \(\mach\) and \(\nach\), any differences in the outputs they
%% produce leaks information about \(\component\).

\paragraph*{Instantiating the Framework}

The property framework has many parameters to support broad applicability.
Conceptually, the following steps are needed to instantiate the framework to a specific machine
and code style: (i) define the base machine semantics; (ii) distinguish payload
from enforcement parts (if any) of the machine state; (iii) identify the set of
security-relevant operations and rules required by the code style; (iv) determine
how to label (or otherwise associate) machine instruction with security-relevant
operations as appropriate; (v) specify the form of observable events.

In \cref{sec:example} we give such an instantiation, and use it to explain the
security properties.  But we stress that the property definitions themselves
are \emph{independent} of the parameter instantiation.

\apt{Suggest moving threat model to here.}

@@

Finally, the core property definitions are given in \cref{tab:props},
arranged to show their commonalities and distinctions. Each definition gives a criterion
quantified over states \(s\) that immediately follow call steps.
If an execution includes a transition \(s' \stepstounder{\overline{\psi}} s\)
where \(\mathbf{call} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then \(s\) is the target
of a call.
%Likewise, if \(\mathbf{tailcall} ~ \addr ~ \overline{\reg} \in \bar{\psi}\), then
%\(s\) is the target of a tailcall.
As a shorthand, we write that each property is defined
by a criterion that must hold ``for all call targets \(s\),'' or, in the case of \(\wbcf\),
``for all call steps \(s \stepstounder{} s'\).''

@@

The operation rule for a tailcall is similar to that for a normal call.
We do not push the caller's view onto the stack,
but replace it outright. This means that a tailcall does not increase the size of
the call stack, and therefore for purposes of our properties, all tailcalls will
be considered to return simultaneously when the eventual {\bf return} operation
pops the top of the stack.

Since the caller will not be returned to, it does not need integrity, but
it should still enjoy confidentiality. We set its frame to \(\unsealed\) rather
than \(\sealed\) to express this. In terms of \cref{tab:props}, we replace
``call'' with ``call or tailcall'' in \(\clrc\), \(\clec\), and \(\clei\).

@@

The centrality of the functional abstraction and its security are behind the
many software and hardware mechanisms proposed for its protection
\cite{Cowan+98, NagarakatteZMZ09, NagarakatteZMZ10, DeviettiBMZ08,
Kuznetsov+14, Dang+15, Shanbhogue+19, Woodruff+14, Chisnall+15,
SkorstengaardLocal, SkorstengaardSTKJFP, Georges22:TempsDesCerises,
DBLP:conf/sp/RoesslerD18, Gollapudi+23}, which we survey briefly in
\cref{sec:intro}. Among these techniques, the micro-policies of Roessler and
DeHon \cite{DBLP:conf/sp/RoesslerD18} are the most flexible in supporting a
range of safeguards including lazy enforcement, and suitable for formulating
and validating strong security guarantees.
